Robust Control of a Fluxonium Qubit

### 2020-11-24 paper/1introduction.tex
progress on todos

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate decoherence
and errors due to parameter deviations
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
decoherence due to longitudinal relaxation and pure dephasing
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

@@

This paper is organized as follows. \todo{use section labels?}
First, we introduce the ALTRO method in the context of QOC.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method. Then, we
outline a method for making the optimization aware of longitudinal
relaxation. Next, we outline three methods for achieving
robustness to static parameter deviations. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise.

>>>>

Analytic and numerical techniques for QOC have been remarkably successful in
designing high fidelity gates (citations needed). 
While these methods are able to predict system behavior with high accuracy, they tend 
to be sensitive to experimental errors such as parameter drift, noise, and finite control 
resolution. 
% In addition to the high
% simulated fidelities which can be obtained with these techniques, it is desirable for the
% resultant quantum state trajectory to be insensitive to experimental errors
% such as parameter drift, noise, and finite control resolution. Quantum state
% trajectories that are insensitive to these effects are said to be robust.
% In line with this objective, 
In order to address these issues,
recent work has sought to mitigate
errors due to parameter deviations.
Considering the dynamical and geometric phases of quantum state
trajectories has led to analytic methods for mitigating
errors due to parameter deviations and pure dephasing
\cite{han2020experimental, merrill2014progress, xu2020nonadiabatic, zhang2020universal}.
Floquet techniques have been experimentally demonstrated to simultaneously mitigate
longitudinal relaxation and pure dephasing decoherence
\cite{huang2020engineering, mundada2020floquet}.
Numerical QOC techniques have been adapted to mitigate longitudinal relaxation
by modeling master equations \cite{rembold2020introduction} and employing
Monte Carlo style quantum trajectories \cite{abdelhafez2019gradient}.
Additionally, efforts have been made to incorporate experimental feedback
into optimization \cite{huang2020engineering}. Comprehensive tools that interleave characterization
and pulse design are under active development \cite{wittler2020integrated}.

@@

This paper is organized as follows.
First, we introduce the ALTRO method in the context of QOC \ref{sec:background}.
We describe realistic constraints for the fluxonium and
map them to the ALTRO method \ref{sec:fluxonium}. Then, we
outline a method for making the optimization aware of longitudinal
relaxation \ref{sec:longitude}. Next, we outline three methods for achieving
robustness to static parameter deviations \ref{sec:static}. Finally,
we employ the robust control techniques to mitigate 1/$f$ flux noise
\ref{sec:stochastic}.

### 2020-11-24 paper/2background.tex
progress on todos

\section{QOC + AL-iLQR}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE)
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional $\nabla_{u} J(u)$.
This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
This scheme lends itself to projected gradient methods
which restrict the optimization to the constraint manifold
\cite{clarkson2010coresets, hauswirth2016projected}.

>>>>

\section{QOC + AL-iLQR \label{sec:background}}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE)
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
%% The evolution is sometimes cast with the evolution
%% of a density matrix under the Lindblad master equation to
%% model the decoherence of the state explicitly.
The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u(t))} \rvert}^{2}$.
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional $\nabla_{u} J(u)$.
This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
This scheme lends itself to projected gradient methods
which restrict the optimization to the constraint manifold
\cite{clarkson2010coresets, hauswirth2016projected}.

### 2020-11-24 paper/3fluxonium.tex
progress on todos

\section{QOC on the Fluxonium}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

>>>>

\section{QOC on the Fluxonium \label{sec:fluxonium}}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

The optimization problem takes the form:
\begin{mini!}[2]
    {x_{0:N}, u_{0:N\text{-}1}}{\sum_{k=1}^N \norm{x_k\text{-}x_f}_{Q_k} + \sum_{k=1}^{N-1} \norm{u_k}_{R_k}}{}{}
    \addConstraint{x_{k+1}}{= f(x_k, u_k)}  \label{eq:dyn_con}
    \addConstraint{x_1}{= x_\text{init}} \label{eq:init_con}
    \addConstraint{x_N}{= x_f} \label{eq:term_con}
\end{mini!}
where the augmented state and controls are given by:
\begin{equation}
  x = \begin{bmatrix} \psi^{(1)} \\ \psi^{(2)} \\ \int a \ dt\\ a \\ \pdv*{a}{t} \end{bmatrix} \quad
  u = \begin{bmatrix} \pdv*[2]{a}{t} \end{bmatrix}
\end{equation}
The initial and final conditions are determined by the
desired unitary transformation, which requires two quantum states
to specify $\psi^{(1)}$, $\psi^{(2)}$.
The ALTRO implementation we use does not currently support complex numbers so
we compute in the isomorphism $\mathcal{H}(\mathbb{R}^{2n}) \cong \mathcal{H}(\mathbb{C}^{n})$
given in \cite{leung2017speedup}:
\begin{equation}
  H \psi \cong \begin{bmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{bmatrix}
  \begin{bmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{bmatrix}
\end{equation}
The dynamics function $f(x_{k}, u_{k})$ integrates the TDSE dynamics for the quantum states
\eqref{eq:tdse}, \eqref{eq:hamiltonian}
and integrates the moments of the flux drive amplitude. Exposing lower order moments of the flux
drive amplitude allows us to penalize their norms, smoothing the flux drive amplitude
and mitigating AWG ringing due to high frequency transitions.
The matrices $Q_{k}$ and $R_{k}$ define the penalty metric. We choose $Q$
to be diagonal because it is computationally efficient. This corresponds
to penalizing phase differences between the quantum states, although
other phase-insensitive metrics such as infidelity may be employed.
The final state transfer is the most important objective which is
encoded in the realtive metric weight $Q_{N} \sim N \cdot Q_{k}$.
%%, $k \in \{1, ..., N - 1\}$.

Additionally, we impose constraints that reflect limitations
of the apparatus and ones that improve the experimental realization
of the control pulse. To ensure gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\psi_{N} = \psi_{f}$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint function
takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum states
are constrained to $1$ to ensure discretization error is not
exploited $c_{k} = x_{k}^{T}x_{k} - 1$.

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude \ref{sec:longitude} and employ robust control techniques to mitigate
pure dephasing \ref{sec:static}, \ref{sec:stochastic}.

### 2020-11-24 paper/4longitude.tex
progress on todos

\section{Longitudinal Relaxation Awareness}
The longitudinal relaxation time $T_{1}$ varies with
control parameters in a range of superconducting circuit platforms.
It is advantageous to tune the controls to extend the longitudinal
relaxation time.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
We avoid this increase in computational complexity by
penalizing the integrated rate of longitudinal relaxation,
i.e. the probability of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation
increase monotonically in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

>>>>

\section{Longitudinal Relaxation Awareness \label{sec:longitude}}
The longitudinal relaxation time $T_{1}$ varies with
control parameters in a range of superconducting circuit platforms.
It is advantageous to tune the controls to extend the longitudinal
relaxation time.
Computing the gate error due to longitudinal relaxation
requires propagating density matrices of size $n \times n$ under master equation
dynamics, rather than state vectors of size $n$ under the TDSE dynamics.
We avoid this increase in computational complexity by
penalizing the integrated rate of longitudinal relaxation,
i.e. the probability of longitudinal relaxation.
Using this probability as proxy for the gate error incurred
is reasonable because losses due to longitudinal relaxation
increase monotonically in time.
This technique can be extended to
error channels which share the monotonically increasing property.
Additionally, for a constant $T_{1}$ time, a shorter gate duration
would favor a lower longitudinal relaxation probability. We allow
the optimizer to tune the gate duration in order to minimize the
longitudinal relaxation probability. Our scheme for time-optimal
control is applicable to any time-optimal problem, not only
the one we study here.

### 2020-11-24 paper/6stochastic.tex
brian edits
progress on todos

\section{Robustness to Stochastic Parameter Deviations}
An additional source of experimental error arises from stochastic, time-varying
Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Magnetic flux noise
modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
It is well studied that the spectral density of $\delta a$ follows a
1/$f$ distribution for a range of devices, consisting primarily of low frequency
noise (citations needed). Analytic methods to combat flux noise
take advantage of the low frequency characteristic and
treat the noise as quasi-static, performing generalizations of the spin-echo technique
to compensate for erroneous drift. This is the strategy employed by the analytic gate
considered here.

We compare the analytic gate and those produced by
the numerical methods discussed in the previous section
on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
The flux noise is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}
\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its sampled deviations
follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
in the augmented state vector. In principle the basic sampling method could be modified
similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
methods require no modification from the static case.

>>>>

\section{Robustness to Stochastic Parameter Deviations \label{sec:stochastic}}
An additional source of experimental error arises from stochastic, time-varying
Hamiltonian parameter deviations. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Magnetic flux noise
modifies the fluxonium qubit's amplitude from its nominal value by an amount $\delta a$.
It is well studied that the spectral density of $\delta a$ follows a
1/$f$ distribution for a range of devices, consisting primarily of low frequency
noise (citations needed). Analytic methods to combat flux noise
take advantage of the low frequency characteristic and
treat the noise as quasi-static, performing generalizations of the spin-echo technique
to compensate for erroneous drift. This is the strategy employed by the analytic gate
considered here.

We compare the analytic gate and those produced by
the numerical methods discussed in the previous section
on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
The flux noise is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}
\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its sampled deviations
follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
in the augmented state vector. In principle the basic sampling method could be modified
similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
methods require no modification from the static case. \todo{can you provide a little more 
detail on how this noise factors into the dynamics? Are you just adding some white 
noise to $a(t)$? Are you also considering uncertainty in $f_q$ at the same time?}

### 2020-11-24 paper/a2unscented.tex
progress on todos

\section{Unscented Transformation}
\label{appendix:unscented}
In this section we outline the full unscented sampling procedure.
First, we remark that ALTRO does not natively support complex numbers, so our optimization
is performed under an isomorphism 
$\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
\cite{leung2017speedup}. 
\todo{this should probably be mentioned explicitly in the main body of the paper. Also 
make sure you're being consistent on your definition of $n$. The definition here doesn't seem 
to match your definition other places, especially when you start augmenting the dynamics.}
\begin{equation}
  H \psi \cong \begin{pmatrix} H_{\textrm{re}} & -H_{\textrm{im}} \\ H_{\textrm{im}} & H_{\textrm{re}}\end{pmatrix}
  \begin{pmatrix} \psi_{\textrm{re}} \\ \psi_{\textrm{im}}\end{pmatrix}
\end{equation}

The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with
deviant parameters $\lambda \in \mathbb{R}^{d}$ and dynamics $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
The nominal initial quantum state is given by $\psi_{0}$ with an associated
positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{n \times n}$
which describes the uncertainty in the initial state.
$P_{0}$ is typically chosen to be non-zero even if the state preparation error is negligible.
The deviant parameter
is assumed to have zero mean and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the dynamics $f$.

>>>>

\section{Unscented Transformation}
\label{appendix:unscented}
In this section we outline the full unscented sampling procedure.
The unscented sampling method concerns a quantum state $\psi \in \mathbb{R}^{2n}$ with
deviant parameters $\lambda \in \mathbb{R}^{d}$ and dynamics $\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
The nominal initial quantum state is given by $\psi_{0}$ with an associated
positive-definite covariance matrix $P_{0} \in \mathbb{R}_{++}^{n \times n}$
which describes the uncertainty in the initial state.
$P_{0}$ is typically chosen to be non-zero even if the state preparation error is negligible.
The deviant parameter
is assumed to have zero mean and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{R}_{++}^{d \times d}$ at knot point $k$. The zero mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the dynamics $f$.
